---
title: "Bayesian Hierarchical Multivariate Regression for Ordinal Data"
author: "Peter Hoff"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Bayesian Hierarchical Multivariate Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides functions to approximate the posterior 
distribution of the parameters in the hierarchical multivariate
regression  model given by
\[
\begin{align*} 
y_{i,j,k} & = f_j( z_{i,k} )  \\
Z_i & = B_i X_i + V_E^{1/2} E_i,   \\
\text{vec}(B_i) & = \text{vec}(B_0)  + V_0^{1/2} e_i 
\end{align*}
\]
where $f_j$ is an (unknown) nondecreasing function for each $j$ and 

- $Z_i\in \mathbb R^{m\times n}$; 
- $X_i\in \mathbb R^{p\times n}$; 
- $B_i\in \mathbb R^{m\times p}$. 

for $i=1,\ldots,r$. 
The error terms $E_i$ and $e_i$ 
consist of independent standard normal entries, and 
are also independent 
across values of $i$. 

The functions $f_1,\ldots,f_p$ are treated semiparametrically using the 
extended rank likelihood. As a result, variable-level 
intercepts and scales are not identifiable. 


### load package
```{r}
library(hmvregression)
```


### simulate some parameters and data
```{r} 
set.seed(1) 

m<-5 ; p<-3 ; n<-25 ; r<-100

B0.0<-rsan(c(m,p))
V0.0<-crossprod( rsan(c(m*p,m*p))  )
VE.0<-crossprod( rsan(c(m,m)) )

XA<-rsan(c(r,p,n))

# latent data
ZA.0<-array(dim=c(r,m,n))
BA.0<-array(dim=c(r,m,p))
for(i in 1:r)
{
  BA.0[i,,]<-B0.0 + matrix( t(chol(V0.0))%*%rnorm(m*p), m,p )
  ZA.0[i,,]<-BA.0[i,,]%*%XA[i,,] + t(chol(VE.0))%*%rsan(c(m,n))
} 

# ordinal data
YA<-ZA.0
for(j in 1:m)
{
  zj<-ZA.0[,j,]
  nj<-1+rpois(1,3)
  fj<-rnorm(nj)*sd(zj)
  yj<-matrix( apply(outer( c(zj), fj, ">"),1,sum) , nrow(zj), ncol(zj) )
  YA[,j,]<-yj
}
```


### starting values 
```{r}
ZA<-array(dim=dim(YA)) ; for(j in 1:m) { ZA[,j,]<-zscores(YA[,j,]) }

BA<-array(dim=c(r,m,p))
EA<-array(dim=c(r,m,n))
for(i in 1:r)
{
  BA[i,,]<-tcrossprod(ZA[i,,],XA[i,,]) %*% solve(tcrossprod(XA[i,,]))
  EA[i,,]<-YA[i,,] - BA[i,,]%*%XA[i,,]
}

B0<-apply(BA,c(2,3),mean)
V0<-var( t(apply(BA,1,c)) )
VE<-var(   apply(EA,2,c) )
```

### check preliminary estimates
```{r}
par(mfrow=c(2,2),mar=c(3,3,1,1))
plot(cov2cor(V0),cov2cor(V0.0)) ; abline(0,1)
plot(cov2cor(VE),cov2cor(VE.0)) ; abline(0,1)
plot(BA/sqrt(sum(BA^2)),BA.0/sqrt(sum(BA.0^2))) ; abline(0,1)
plot(B0/sqrt(sum(B0^2)),B0.0/sqrt(sum(B0.0^2))) ; abline(0,1)
```

### MCMC
```{r,fig.keep='last',results='hide'}

B0PSUM<-matrix(0,m,p) 
B0PSAM<-NULL
KT<-NULL

for(s in 1:1000)
{

  BA<-rBA_fc(ZA,XA,B0,V0,VE)

  VE<-rVE_fc(ZA,XA,BA)

  B0<-rB0_fc(BA,V0)

  V0<-rV0_fc(BA,B0)

  # update latent data
  EZA<-ZA ; for(i in 1:r){ EZA[i,,]<-BA[i,,]%*%XA[i,,] }
  for(j in 1:m)
  {
    if(m>1)
    {
      szj<-sqrt(c( VE[j,j]-VE[j,-j]%*%solve(VE[-j,-j])%*%VE[-j,j]  ))
      ezj<-EZA[,j,] +
           apply( ZA[,-j,]-EZA[,-j,],c(1,3),"%*%",solve(VE[-j,-j])%*%VE[-j,j] )
    }
    if(m==1){ szj<-sqrt(c(VE)) ; ezj<-EZA[,1,] }
    ZA[,j,]<-rz_fc(YA[,j,],ZA[,j,],ezj,szj)
  }

  # output 
  if(s%%10==0) 
  {
    B0PSAM<-rbind(B0PSAM,c(B0/sqrt(sum(B0^2)))) 
    B0PSUM<-B0PSUM+10*B0/sqrt(sum(B0^2))

    kt<-NULL
    for(j in 1:m){ kt<-c(kt,kendalltau(ZA.0[,j,],ZA[,j,]) ) }
    KT<-rbind(KT,kt) 

    par(mfrow=c(4,2),mar=c(3,3,1,1))
    plot(cov2cor(V0)[upper.tri(V0)],cov2cor(V0.0)[upper.tri(V0)]) ; abline(0,1)
    plot(cov2cor(VE)[upper.tri(VE)],cov2cor(VE.0)[upper.tri(VE)]) ; abline(0,1)
  
    plot(BA/sqrt(sum(BA^2)),BA.0/sqrt(sum(BA.0^2))) ; abline(0,1)
    plot(B0/sqrt(sum(B0^2)),B0.0/sqrt(sum(B0.0^2))) ; abline(0,1)
 
    matplot(B0PSAM,type="l",lty=2,col=1:prod(dim(B0))) 
    abline(h=c(B0.0/sqrt(sum(B0.0^2))),col=1:prod(dim(B0))) 
    plot(B0.0/sqrt(sum(B0.0^2)),B0PSUM/s) ; abline(0,1) 

    matplot(KT,type="l",col=1:m,lty=2) 
    abline(h=apply(KT,2,mean),col=1:m) 
  
   cat( mean( (ZA-EZA)^2 ),"\n") 
  }
}
```

Keep in mind that 
variable-level intercepts and scales are not identifiable. 


