---
title: "Bayesian Hierarchical Multivariate Regression"
author: "Peter Hoff"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Bayesian Hierarchical Multivariate Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides functions to approximate the posterior 
distribution of the parameters in the hierarchical multivariate
regression  model given by
\[
\begin{align*} 
Y_i & = B_i X_i + V_E^{1/2} E_i,   \\
\text{vec}(B_i) & = \text{vec}(B_0)  + V_0^{1/2} e_i 
\end{align*}
\]
where 

- $Y_i\in \mathbb R^{m\times n}$; 
- $X_i\in \mathbb R^{p\times n}$; 
- $B_i\in \mathbb R^{m\times p}$. 

for $i=1,\ldots,r$. 
The error terms $E_i$ and $e_i$ 
consist of independent standard normal entries, and 
are also independent 
across values of $i$. 



### load package
```{r}
library(hmvregression)
```


### simulate some parameters and data
```{r} 
m<-5 ; p<-3 ; n<-25 ; r<-100

B0.0<-rsan(c(m,p))
V0.0<-crossprod( rsan(c(m*p,m*p))  )
VE.0<-crossprod( rsan(c(m,m)) )

XA<-rsan(c(r,p,n))

YA<-array(dim=c(r,m,n))
BA.0<-array(dim=c(r,m,p))
for(i in 1:r)
{
  BA.0[i,,]<-B0.0 + matrix( t(chol(V0.0))%*%rnorm(m*p), m,p )
  YA[i,,]<-BA.0[i,,]%*%XA[i,,] + t(chol(VE.0))%*%rsan(c(m,n))
}
```


### starting values 
```{r}
BA<-array(dim=c(r,m,p))
EA<-array(dim=c(r,m,n))
for(i in 1:r)
{
  BA[i,,]<-tcrossprod(YA[i,,],XA[i,,]) %*% solve(tcrossprod(XA[i,,]))
  EA[i,,]<-YA[i,,] - BA[i,,]%*%XA[i,,]
}

B0<-apply(BA,c(2,3),mean)
V0<-var( t(apply(BA,1,c)) )
VE<-var(   apply(EA,2,c) )
```

### check preliminary estimates
```{r}
par(mfrow=c(2,2),mar=c(3,3,1,1))
plot(V0,V0.0) ; abline(0,1)
plot(VE,VE.0) ; abline(0,1)
plot(BA,BA.0) ; abline(0,1)
plot(B0,B0.0) ; abline(0,1)
```

### MCMC
```{r,fig.keep='last'}

B0PSUM<-matrix(0,m,p) 
B0PSAM<-NULL

for(s in 1:100)
{

  BA<-rBA_fc(YA,XA,B0,V0,VE)

  VE<-rVE_fc(YA,XA,BA)

  B0<-rB0_fc(BA,V0)

  V0<-rV0_fc(BA,B0)

  # output 
  B0PSAM<-rbind(B0PSAM,c(B0)) ; B0PSUM<-B0PSUM+B0

  par(mfrow=c(3,2),mar=c(3,3,1,1))
  plot(V0,V0.0) ; abline(0,1)
  plot(VE,VE.0) ; abline(0,1)
  plot(BA,BA.0) ; abline(0,1)
  plot(B0,B0.0) ; abline(0,1)
  matplot(B0PSAM,type="l",lty=2,col=1:prod(dim(B0))) 
  abline(h=c(B0.0),col=1:prod(dim(B0))) 
  plot(B0.0,B0PSUM/s) ; abline(0,1) 

}
```




